---
title: "Rain Movies"
author: "Matthew Ross"
date: "11/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Using R to make movies

While most of the visualizations made in R are `static`, with a single image 
displaying the data, one can also make videos in R. These are particularly 
useful when data has a temporal component or a third axis for display. Here
we will show a few ways to make videos using R and airport weather data 

```{r setup, include=FALSE}
library(tidyverse)
library(riem) # Precipitation data
library(animation) #movie making
library(USAboundaries) # boundary data
library(gganimate) #another way to make movies
library(lubridate)
library(ggthemes)
library(tmap)
library(sf)
library(gifski)
library(tigris)
library(elevatr)
theme_set(theme_few()) # Set ggplot theme


knitr::opts_chunk$set(echo = TRUE)
```


# Data acquisition and reshaping


## Finding our station names

```{r}
#riem_networks() # look for network names

co_stations <- riem_stations(network = 'CO_ASOS')

#View(co_stations)

stations <- c('AFF', 'FNL')

```

## Download station data

```{r}

af_p <- riem_measures(station='AFF',date_start = '2018-10-01',
                      date_end = '2019-09-30')

fc <- riem_measures(station='FNL',date_start = '2018-10-01',
                      date_end = '2019-09-30')

View(fc)

```


## Convert both to daily

```{r}

daily_p <- fc %>% 
  bind_rows(af_p) %>%
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()

```


# Making videos 

## Boring first video

```{r}

slice_p <- daily_p %>%
  slice(1:20)

p1 <- ggplot(slice_p,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_point() +
  labs(title = 'Date: {frame_time}', x='Date',y= 'Mean Temp (f)') + 
  transition_time(date) 


animate(p1, nframes=nrow(slice_p))


```


## Leaving a trace

```{r}
p2 <- ggplot(daily_p,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_line() + 
  geom_point() +
  labs(title = '100 mile temp diff', x='Date',y= 'Mean Temp (f)') + 
  transition_reveal(date) + 
  scale_color_few()


animate(p2, nframes=nrow(daily_p))



```


## Static background plot


```{r}
usa <- us_states()

co <- usa %>%
  filter(name == 'Colorado')

sf_p <- daily_p %>%
  st_as_sf(coords=c('lon','lat'),crs=4326)


co_elev <- elevatr::get_elev_raster(co,z=6)

tm_shape(co) + 
  tm_polygons() + 
  tm_shape(co_elev) +
  tm_raster(style='cont',palette = 'Greens') + 
  tm_shape(sf_p) + 
  tm_dots(col='station',size=3,palette=c('red','white'))

```



## Quick plot

```{r}
  
basemap <- tm_shape(co) + 
    tm_polygons() + 
  tm_shape(co_elev) +
    tm_raster(style='cont',
              palette = 'Greens',
              title='Elevation (m)') 
```



```{r,eval=F}
 
full_dates <- unique(sf_p$date)

sequence <- 1:length(full_dates)
saveGIF({
  for(i in sequence){
  sub_sf <- sf_p %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(col='station',
            size=2,
            palette = c('red','white'),
            legend.show = F) 
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i]) + 
          tm_add_legend(type='symbol',
                        labels=c('Air Force',
                                 'Fort Collins'),
                        col=c('red','white'),
                        title='Station'))
  }
},movie.name = 'precip_flash.gif',
  interval=0.05,
  ani.width=1000,
  ani.height=800)


```

# A cooler video

## Let's see if we can see weather fronts by downloading all Colorado data and
trying to get those to flash


```{r,eval=F}

all_co <- map(co_stations$id,~riem_measures(.x,date_start='2018-10-01',
                                         date_end = '2019-09-30'))
stack_co <- do.call('rbind',all_co)

daily_p <- stack_co %>% 
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()

save(daily_p,file='DailyP.RData')
```

## Full state video

This code is almost identical to the above, but it uses all ASOS/AWOS precip
stations


```{r, eval=F}
load('DailyP.RData')

daily_sf <- daily_p %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326,remove=F)




full_dates <- unique(daily_sf$date)

sequence <- 1:length(full_dates)
saveGIF({
  for(i in sequence){
  sub_sf <- daily_sf %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(title.size='Daily Precip (in)',
            col='lon',
            size='daily_p',
            style='cont',
            palette='Oranges',
            legend.show = F,
            size.max=19,
            scale=3.5,
            sizes.legend=c(0,1,5,10,20))
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i],
        legend.position=c(0.7,0.2)))
  }
},movie.name = 'precip_flash_all.gif',
  interval=0.12,
  ani.width=1000,
  ani.height=800)



```



# In Class work.

## GGanimate

Make a video showing the correlation between Air Force Academy and Fort Collins
between the min, mean, and max temps. The movie should cover the full year. 

```{r}
filtered_data <- daily_p %>%
  filter(station %in% c("AFF", "FNL"))

plot1 <- ggplot(filtered_data) +
  geom_point(aes(x=date,y=min_temp,color=station, group = seq_along(min_temp), shape = 0)) + 
  geom_point(aes(x = date, y = max_temp, color = station, group = seq_along(max_temp), shape = 3)) +
  geom_point(aes(x = date, y = mean_temp, color = station, group = seq_along(mean_temp))) +
  labs(x='Date',y= 'Temp (f)')+
  transition_reveal(date)
 

animate(plot1, nframes=nrow(filtered_data))

change legend, size or shape of min, mean, and max temps to show difference

```

#spread by station

## (Animation package)

Make a video of mean temperature differences for the full Colorado dataset
(you need to download this [here](https://www.dropbox.com/s/s7zt18elgoh8q2b/DailyP.RData?dl=0)) first.

The video should look similar to the one above, but I want you to choose your own 
style (change the background, add other layers (like counties or cities)), etc...

Creating a basemap for the final product
```{r}
us <- us_states()

co <- us %>%
  filter(name == 'Colorado')

elevation <- elevatr::get_elev_raster(co, z=6)

counties_co <- counties(state = "Colorado", cb = TRUE)

elev_co_crop <- crop(elevation, counties_co)

reproj_co <- spTransform(counties_co, crs(elevation))

base <- tm_shape(elev_co_crop)+
  tm_raster(style= 'cont',
            palette = 'Pinks',
            title = "elevation (m)")+
  tm_borders("black")
  tm_shape(reproj_co)
```


```{r}
mean_temp <- daily_p %>%
  dplyr::select(1:5)

slice_mean <- mean_temp %>%
  slice(1:100)

slice_sf <- slice_mean %>%
  st_as_sf(., coords = c('lon', 'lat'), crs = 4326, remove = F)

all_dates_slice <- unique(slice_sf$date)

sequence <- 1:length(all_dates_slice)
saveGIF({
  for(i in sequence){
    small_sf <- slice_sf %>%
      filter(data == all_dates_slice[i])
    
    my_map <- base + tm_shape(subset_sf)+
      tm_sots(title.size = 'Mean Temperature (ºC)',
              col = 'mean_temp',
              size = 'mean_temp',
              scale = 1.5,
              style = 'cont',
              pallate = 'Blues')
    print(my_map +
            tm_layout(title=all_dates_slice[i],
                 legend.position = c(0.8, 0.3)))
    
  }
}, movie.name = 'mean_temp_flashes.gif',
  interval = 0.12,
  ani.width = 1000,
  ani.height = 800)
```


